---
title: "Reproducible research"
author: "Toghrul Abilli"
date: "10/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is R markdown for explotary data analysis

1. Code for reading in the dataset and/or processing the data

```{r setwd, echo = FALSE}

setwd('C:/Users/tabilli/Desktop/Projects/Rproject/Reproducible')
```

After setting working directory, csv data read from source.
```{r readdata, echo=TRUE}

activity<-read.csv('activity.csv')

```
To understand data, we should look at summaries.
As shown from the table, many NA's are observed in step column. 
```{r summary, echo=TRUE}

summary(activity)

```
Lets look at size of data using dim function.
```{r dim, echo=TRUE}

dim(activity)

```

2. Histogram of the total number of steps taken each day

```{r hist, echo=TRUE}
steps_each_day<-aggregate(steps ~ date, activity, sum)
hist(steps_each_day$steps, col = "blue")

```

3. Mean and median number of steps taken each day

```{r median, echo=TRUE}
median(steps_each_day$steps)
mean(steps_each_day$steps)
```

4. Time series plot of the average number of steps taken

```{r time, echo=TRUE}
stepsbyinterval<-aggregate(steps ~ interval, activity, mean)
plot(stepsbyinterval$interval, stepsbyinterval$steps, type = "l", xlab = "Interval", ylab = "Average steps",main = "Average number of steps")
```

5. The 5-minute interval that, on average, contains the maximum number of steps

```{r maxinterval, echo=TRUE}
max_interval<-stepsbyinterval[which.max(stepsbyinterval$steps),1]
max_interval

```

6. Code to describe and show a strategy for imputing missing data


```{r impute, echo=TRUE}

activity$steps[is.na(activity$steps)]<-mean(activity$steps,na.rm=TRUE)

```

7. Histogram of the total number of steps taken each day after missing values are imputed

```{r hist1, echo=TRUE}
stepsbyday_1<-aggregate(steps ~ date, activity, sum)

hist(stepsbyday_1$steps, col="blue",xlab="Steps", ylab = "Frequency", main = "Total number of steps taken each day after missing values", breaks = 10)
```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r plotcompare, echo=TRUE}

weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday",  "Friday")
activity$dtype = as.factor(ifelse(is.element(weekdays(as.Date(activity$date)),weekdays), "Weekday", "Weekend"))
steps_by_interval_i <- aggregate(steps ~ interval + dtype, activity, mean)
library(lattice)

xyplot(steps_by_interval_i$steps ~ steps_by_interval_i$interval|steps_by_interval_i$dtype, main="Average number of steps per interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```





